Time Series Analysis of Chaotic Dynamical Systems with Recurrent Neural Networks

### 2019-04-17 doc/paper/paper.tex
line formattings in paper.tex

An example of this is the trajectories of the sun and planets in our solar
system. Newton's equations can be used to create a system of equations to
model the bodies of mass (known as the n-body problem). However, with the
existence of other bodies of mass from other solar systems within the
universe, small perturbations are introduced into the system. If the system
is chaotic, then over time, the actual trajectories will diverge from our
model's \cite{oestreicher2007history}.

One of the defining characteristics of chaos is that the error from
perturbations grows exponentially \cite{oestreicher2007history}. At the time
of Lorenz's paper, unstable error was not well understood, and this led to
Lorenz's surprising discovery \cite{oestreicher2007history}. Lorenz reported
that a single iteration in calculating the solution to the dynamical system
took approximately one second \cite{lorenz1963deterministic}. In his paper,
he shows one of his calculations having about 3000 iterations, which must
have taken around 50 minutes. Given the amount of time it takes to perform
the calculation, it's tempting to continue the calculation from the output of
the previous calculation. However, the story of the discover is that while
the computer used 6 digit accuracy, it only output 3 digits, so when the
simulation was continued by Lorenz with the less accurate measurements, he
found the results to be very different \cite{oestreicher2007history}.

@@

\subsection{Calculating the Largest Lyapunov Characteristic Exponent}

\label{sec:calc_lya}
In order to find the Lyapunov time, we only need to solve for the largest
LCE. To find this, we will use the following method detailed in
\cite{viswanath1998lyapunov}.

@@

\subsection{Recurrent Neural Networks and Traditional Training Method}
\label{sec:rnn}
RNNs are usually designed by having $K$ input units, $L$ output units, 
and some internal units organized by some network structure. These units
have evolving values for every time point. The
structure of networks describes how units are connected. 
The input units at time $t_i$ are set to $\uv(t_i)$. However, output
units and
internal units are calculated from the values of other units. These other 
units are said to be connected to the resultant unit. Output
units are calculated from current input units, current internal units,
and from previous output units. Internal units 
are calculated from the values of the current input units, 
previous internal units, and previous output units.

>>>>

An example of this is the trajectories of the sun and planets in our solar
system. Newton's equations can be used to create a system of equations to
model the bodies of mass (known as the n-body problem). However, with the
existence of other bodies of mass from other solar systems within the
universe, small perturbations are introduced into the system. If the system
is chaotic, then over time, the actual trajectories will diverge from our
model's \cite{oestreicher2007history}.
One of the defining characteristics of chaos is that the error from
perturbations grows exponentially \cite{oestreicher2007history}. At the time
of Lorenz's paper, unstable error was not well understood, and this led to
Lorenz's surprising discovery \cite{oestreicher2007history}. Lorenz reported
that a single iteration in calculating the solution to the dynamical system
took approximately one second \cite{lorenz1963deterministic}. In his paper,
he shows one of his calculations having about 3000 iterations, which must
have taken around 50 minutes. Given the amount of time it takes to perform
the calculation, it's tempting to continue the calculation from the output of
the previous calculation. However, the story of the discover is that while
the computer used 6 digit accuracy, it only output 3 digits, so when the
simulation was continued by Lorenz with the less accurate measurements, he
found the results to be very different \cite{oestreicher2007history}.

@@

\subsection{Calculating the Largest Lyapunov Characteristic Exponent}
\label{sec:calc_lya}

In order to find the Lyapunov time, we only need to solve for the largest
LCE. To find this, we will use the following method detailed in
\cite{viswanath1998lyapunov}.

@@

\subsection{Recurrent Neural Networks and Traditional Training Method}
\label{sec:rnn}

RNNs are usually designed by having $K$ input units, $L$ output units, and
some internal units organized by some network structure. These units have
evolving values for every time point. The structure of networks describes how
units are connected. The input units at time $t_i$ are set to $\uv(t_i)$.
However, output units and internal units are calculated from the values of
other units. These other units are said to be connected to the resultant
unit. Output units are calculated from current input units, current internal
units, and from previous output units. Internal units are calculated from the
values of the current input units, previous internal units, and previous
output units.
